Assessing landscape N removal in coastal New England catchments using the

### 2023-10-26 manuscript/manuscript.Rmd
local changes

---
title: "Assessing landscape N removal in coastal New England catchments using the
  N-Sink approach with a new R package, nsink"
author:
- "Kellogg, D. Q.^\\*^^*1*^, Hollister. J. W. ^*2*^, Arnold, C. L.^*3*^, Gold, A.
  J. ^*1*^, Wilson, E. H.^*3*^, Chadwick, C.^*3*^, Dickson, D.^*3*^, Lei-Parent,
  Qian^*3*^, Forshay, K. J.^*4*^"
- ^*1*^ University of Rhode Island, Department of Natural Resources Science, Kingston,
  RI 02881
- ^*2*^ US Environmental Protection Agency, Office Of Research and Development, Atlantic
  Coastal Environmental Sciences Division, Narragansett, RI 02882
- ^*3*^ University of Connecticut, Center for Land Use Education and Research, Haddam,
  CT 06438
- ^*4*^ US Environmental Protection Agency, Office Of Research and Development, Groundwater
  Characterization and Remediation Division, Ada, OK 74820
- '^\*^corresponding author: <qkellogg@uri.edu>'
bibliography: manuscript.bib
output:
  word_document:
    keep_md: true
    reference_docx: manuscript_template.docx
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
csl: ecology.csl
abstract: "Excess nitrogen (N) loading to coastal ecosystems impairs estuarine water
  quality. Land management decisions made within estuarine watersheds have a direct
  impact on downstream N delivery. Natural features frequently found within the headwaters
  of watersheds can act as landscape sinks for N, such as wetlands, streams and ponds
  that transform dissolved N into gaseous N, effectively removing it from the aquatic
  system. Identifying and evaluating these landscape sinks and their spatial relationship
  to N sources can help managers understand the effects of alternative decisions on
  downstream resources. N-Sink is an approach that uses widely available GIS data
  to identify landscape sinks within HUC-12 (or larger) catchments, estimate their
  N removal potential and map the effect of those sinks on N movement through the
  catchment. A series of static maps are produced to visualize N removal efficiency,
  transport and delivery, the latter in the form of an index. To facilitate data acquisition,
  processing and visualization, an R package `nsink` was developed and applied to
  produce static maps for all HUC-12 catchments in southern New England. The R package
  can be used to develop N delivery maps for a specific HUC-12 of interest, or users
  can visit the University of Connecticut website to explore previously mapped areas.
  Users can also investigate specific flowpaths interactively by clicking on any location
  within the catchment. A flowpath is generated, along with a table describing N removal
  along each segment."
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  tidy = FALSE,
  message = FALSE,
  fig.height = 4.8,
  fig.width = 6,
  dpi = 300
)
#load(system.file("testdata.rda", package = "nsink"))
#load("testdata.rda")
```

@@

```{r get-data, eval = TRUE, cache=FALSE, include=FALSE}
# eval = TRUE if you want to download the data
# Get data for selected HUC
nsink::nsink_get_data(niantic_huc_id, data_dir = "nsink_niantic_data/")
```

@@

```{r compare-flowpaths, include = TRUE, echo = FALSE}
knitr::kable(niantic_fp_removal_b, caption = "Table 3: Summary of N removal along specified flowpath B.")
```

@@

```{r table4-build, eval = TRUE, include = FALSE}
table3 <- data.frame("Land Cover" = c("Developed", "Agriculture", "Forest/Vegetated",
                                      "Wetlands", "Other"),
                     "West (%)" = c("23","7","64","6","0"),
                     "East (%)" = c("20","0","61","18","1"),
                     check.names = FALSE)
```
```{r table4, include = TRUE, echo = FALSE}
knitr::kable(table3, caption = "Table 4: Summary of land cover in highlighted areas of interest.")
```

@@

An analysis of N delivery from these two similarly developed areas (average N loading index of developed areas: West = 0.57; East = 0.56),  show the western area may contribute more N loading to the estuary than the eastern area, with the average delivery index value of the western area (41) almost 80% higher than the eastern area (23; Figure 5C), suggesting a greater need for N source controls within the western development.

```{r, plot5C, include = TRUE, echo = FALSE, warning = FALSE, fig.cap = "Figure 5C: Nitrogen Delivery Index within an area of interest in the Niantic River HUC-12, highlighting two similarly developed areas. Darker colors denote higher N delivery based on normalized loading from different land covers multiplied by the N Transport Efficiency. "}

>>>>

---
title: "Assessing landscape N removal in coastal New England catchments using the
  N-Sink approach with a new R package, nsink"
author:
- "Kellogg, D. Q. ^*1*^, Hollister. J. W. ^\\*^ ^*2*^, Arnold, C. L.^*3*^, Gold, A.
  J. ^*1*^, Wilson, E. H.^*3*^, Chadwick, C.^*3*^, Dickson, D.*^*3*^, Lei-Parent,
  Qian*^*3*^, Forshay, K. J.^*4*^"
- ^*1*^ University of Rhode Island, Department of Natural Resources Science, Kingston,
  RI 02881
- ^*2*^ US Environmental Protection Agency, Office Of Research and Development, Atlantic
  Coastal Environmental Sciences Division, Narragansett, RI 02882
- ^*3*^ University of Connecticut, Center for Land Use Education and Research, Haddam,
  CT 06438
- ^*4*^ US Environmental Protection Agency, Office Of Research and Development, Groundwater
  Characterization and Remediation Division, Ada, OK 74820
- '^\*^ \*corresponding author: <qkellogg@uri.edu*>'
bibliography: manuscript.bib
output:
  pdf_document: default
  word_document:
    keep_md: true
    reference_docx: manuscript_template.docx
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
csl: ecology.csl
abstract: "Excess nitrogen (N) loading to coastal ecosystems impairs estuarine water
  quality. Land management decisions made within estuarine watersheds have a direct
  impact on downstream N delivery. Natural features frequently found within the headwaters
  of watersheds can act as landscape sinks for N, such as wetlands, streams and ponds
  that transform dissolved N into gaseous N, effectively removing it from the aquatic
  system. Identifying and evaluating these landscape sinks and their spatial relationship
  to N sources can help managers understand the effects of alternative decisions on
  downstream resources. N-Sink is an approach that uses widely available GIS data
  to identify landscape sinks within HUC-12 (or larger) catchments, estimate their
  N removal potential and map the effect of those sinks on N movement through the
  catchment. A series of static maps are produced to visualize N removal efficiency,
  transport and delivery, the latter in the form of an index. To facilitate data acquisition,
  processing and visualization, an R package `nsink` was developed and applied to
  produce static maps for all HUC-12 catchments in southern New England. The R package
  can be used to develop N delivery maps for a specific HUC-12 of interest, or users
  can visit the University of Connecticut website to explore previously mapped areas.
  Users can also investigate specific flowpaths interactively by clicking on any location
  within the catchment. A flowpath is generated, along with a table describing N removal
  along each segment."
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  tidy = FALSE,
  message = FALSE,
  fig.height = 4.8,
  fig.width = 6
)
#load(system.file("testdata.rda", package = "nsink"))
#load("testdata.rda")
```

@@

```{r get-data, eval=TRUE, cache=FALSE, include=FALSE}
# eval = TRUE if you want to download the data
# Get data for selected HUC
niantic_download <- nsink::nsink_get_data(niantic_huc_id, 
                                  data_dir = "nsink_niantic_data/")
```

@@

```{r compare-flowpaths, include = TRUE, echo = FALSE}
knitr::kable(niantic_fp_removal_b, caption = "Table 3: Summary of N removal along specified flowpath B")
```

@@

```{r table4-build, eval = TRUE, include = FALSE}
table3 <- data.frame("Land Cover" = c("Developed", "Agriculture", "Forest/Vegetated",
                                      "Wetlands", "Other"),
                     "West (%)" = c("23","7","64","6","0"),
                     "East (%)" = c("20","0","61","18","1"),
                     check.names = FALSE)
```
```{r table4, include = TRUE, echo = FALSE}
knitr::kable(table3, caption = "Table 3: Summary of Land Cover in Highlighted Areas of Interest")
```

@@

An analysis of N delivery from these two similarly developed areas (average N loading index of developed areas: West = 0.57; East = 0.56),  show the western area may contribute more N loading to the estuary than the eastern area, with the average delivery index value of the western area (41) almost 80% higher than the eastern area (23; Figure 5c), suggesting a greater need for N source controls within the western development.

```{r, plot5c, include = TRUE, echo = FALSE, warning = FALSE, fig.cap = "Figure 5C: Nitrogen Delivery Index within an area of interest in the Niantic River HUC-12, highlighting two similarly developed areas. Darker colors denote higher N delivery based on normalized loading from different land covers multiplied by the N Transport Efficiency. "}
